\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-11-10 coqimpl.tex
edits

The implementation works by translating programs in \Lang syntax into
programs that can be checked and evaluated by Coq.
The translation is compatible with interactive theorem proving, in that
a family is translated piece by piece, allowing each field to be defined
and checked separately.
The translation is modular and efficient, in that
code compiled for non-further-bindable fields of a base family
is shared with derived families without having to be rechecked.

@@

For instance,
consider field \lsti{env} in family \lsti{STLC}.
It is translated into a top-level module named \lsti{STLC\_env}.
This module has a parameter called \lsti{self} representing the enclosing
family: fields of the current family in the context of \lsti{env} can be
referenced through \lsti{self}.
In particular, \lsti{env} is defined as \lsti{id}\,->\,\lsti{option ty},
where \lsti{ty} is a late-bound reference to the \lsti{ty} field of the
enclosing family.
Hence, this reference to \lsti{ty} is translated in \lsti{STLC\_env} to \lsti{self.ty},
which is manifestly polymorphic to the enclosing family.
This translation of the \lsti{env} field can be shared
with a derived family even if it extends \lsti{ty} (e.g., \lsti{STLCProd})---%
no recompilation is needed because \lsti{self.ty} is not tied to a
particular definition of \lsti{ty}.

The type of \lsti{STLC\_env}'s \lsti{self} parameter is \lsti{STLC\_env\_Ctx},
a module type constructed from the translation of \lsti{env}'s preceding
field, \lsti{STLC\_ty}, and its context \lsti{STLC\_ty\_Ctx}.
In turn, \lsti{STLC\_ty\_Ctx} (not shown in \cref{fig:stlc-compiled}) is
constructed from the translation of \lsti{ty}'s preceding field,
\lsti{STLC\_subst}, and its context \lsti{STLC\_subst\_Ctx}.
Thus, the \lsti{self} parameter can be used to reference those and only
those fields in the current field's typing context, which echoes the
discussion in \cref{sec:sound}.

@@

An \lsti{FInductive} definition is translated to a parameterized module type.
Consider the inductive type \lsti{tm}.
In \lsti{STLC}, it is translated to a top-level module type
\lsti{STLC\_tm} that declares a \lsti{tm} type,
its constructors (\lsti{tm_unit} etc.),
its partial recursor (\lsti{tm_prect_STLC}),
and the propositional equalities on the partial recursor (\lsti{tm_unit_eq_STLC} etc.).

@@

The command \lsti{FInductive tm : Set += tm_fix : }\dadada~in family
\lsti{STLCFix} is again translated to a module type \lsti{STLCFix\_tm} (\cref{fig:stlcfix-compiled}).
It includes all names declared by \lsti{STLC\_tm}, and additionally declares
the new constructor \lsti{tm_fix} and related equalities.

@@

Upon \lsti{End}~\lsti{subst}, a module type named \lsti{STLC\_subst} is generated.
%
As discussed in \cref{sec:latebind}, \lsti{subst} can be further bound,
so its definition is not exposed to the succeeding fields.
Accordingly, the translation \lsti{STLC\_subst} merely declares the type of \lsti{subst}
and the existence of its propositional equalities, leaving \lsti{subst} undefined
and the equalities unproven.
%
The propositional equalities are stated in terms of the case handlers,
whose definitions \emph{are} available through the \lsti{self} parameter.
For example, the type of \lsti{subst_tm_unit_eq} can be simplified to
\lsti{\forall x t, self.subst self.tm_unit x t = self.tm_unit}.
The equalities will be available for use in the checking and translation of
the succeeding fields through their \lsti{self} parameters.

Importantly, code generated for the case handlers is shared with derived families.
In \cref{fig:stlcfix-compiled}, \lsti{STLCFix\_subst\_Cases}
reuses---without rechecking---all the case handlers in
\lsti{STLC\_subst\_Cases} via the command \lsti{Include STLC\_subst\_Cases(self)}.

@@

In family \lsti{STLC},
field \lsti{env} and the case handlers for \lsti{subst}
are not further-bindable by derived families.
In contrast, \lsti{tm}, \lsti{subst}, and the related equalities
can be further bound.
The distinction is reflected in the translations.
The further-bindable fields are translated to module types that export only types of the fields.
The non-further-bindable fields are translated to modules that export definitional equalities
on the fields.
Opaque definitions can be further bound, too (\cref{sec:override});
they are translated to modules that export opaque definitions.

@@

Upon the conclusion of a family definition (e.g., \lsti{End}~\lsti{STLCFix}),
\lsti{self} parameters in the translations of the fields must be instantiated,
so that references like \lsti{STLCFix.typesafe} can be translated.
This is done by assembling the \lsti{self}-parameterized pieces into a single module.
%
For example, module \lsti{STLC} in \cref{fig:stlc-compiled} is generated
upon \lsti{End}~\lsti{STLC}.
All the fields are added to this module in the same order as they appear in the family definition.

For the non-further-bindables, the translated modules are directly included
(e.g., \lsti{Include STLC\_subst\_Cases} and \lsti{Include STLC\_env}).
%
The instantiations of \lsti{self}s are implicit thanks to a Coq nicety:
when including a higher-order module, Coq automatically instantiates its
parameter with the current interactive module environment.
For instance, command \lsti{Include STLC\_subst\_Cases} is successfully executed,
because Coq automatically instantiates \lsti{self} using the current
module environment, which contains all the fields required by \lsti{STLC\_subst\_Cases\_Ctx}.

For the further-bindables, \lsti{Axiom}s declared in the module types must be instantiated.
%
\begin{itemize}[align=left,itemsep=0pt,labelsep=*,leftmargin=*]

\item
An inductive type \lsti{tm} is generated, instantiating the axiomatized
\lsti{tm} type and its constructors.
%
The partial recursor \lsti{tm_prect_STLC} is defined
%to \lsti{\lambda}\,\lsti{P, tm_rect (\lambda}\,\lsti{t, option (P t))},
with the help of \lsti{tm_rect}, the recursor Coq generates for \lsti{tm}.
Proving the equalities on \lsti{tm_prect_STLC} is then trivial, by \lsti{reflexivity}.

\item
Similarly, \lsti{subst} is instantiated by applying the recursor \lsti{tm_rect}
to the (already included) case handlers (\lsti{subst\_tm_unit}, etc.).
Propositional equalities on \lsti{subst} are then trivial, by \lsti{reflexivity}.

@@

Module \lsti{STLCFix} in \cref{fig:stlcfix-compiled} is generated upon
\lsti{End}~\lsti{STLCFix}, in the same fashion described above for \lsti{STLC}.
The translation makes code and proof sharing evident.
In particular, case handlers compiled for \lsti{STLC}
are reused to instantiate \lsti{subst}, \lsti{substlem}, and alike.
\lsti{STLC\_env} and \lsti{STLC\_typesafe}
are also reused to construct module \lsti{STLCFix}.
%
One may argue that because the first four constructors of \lsti{tm} are
repeated in \lsti{STLCFix}, the translation does not satisfy the
modular compilation requirement.
We could address this issue by using wrapper types,
but we consider restating constructors a small price to pay in return
for the clarity and concision of implementation.
We emphasize that compiled case handlers
are entirely reusable without rechecking, even with restated constructors.

Finally, the reference \lsti{STLCFix.typesafe} (where \lsti{STLCFix} is a family)
can simply be translated to \lsti{STLCFix.typesafe} (where \lsti{STLCFix} is a Coq module),
as the last line of \cref{fig:stlcfix-compiled} shows.

>>>>

The implementation works by translating programs in \Lang syntax into
programs that can be checked and evaluated by Coq.
The translation is compatible with interactive theorem proving (\ref{chg:tooling}), in that
a family is translated piece by piece, allowing each field to be defined
and checked separately.
The translation is modular and efficient, in that
code compiled for fields of a base family
can be shared with derived families without having to be rechecked.

@@

As an example,
consider field \lsti{env} in family \lsti{STLC}.
It is translated into a top-level module named \lsti{STLC\_env}.
This module has a parameter called \lsti{self} representing the enclosing
family: fields of the current family in the context of \lsti{env} can be
referenced through \lsti{self}.
In particular, \lsti{env} is defined as \lsti{id}\,\lsti{->}\,\lsti{option ty},
where \lsti{ty} is a late-bound reference to the \lsti{ty} field of the
enclosing family.
Hence, this reference to \lsti{ty} is translated to \lsti{self.ty},
which is manifestly polymorphic to the enclosing family.
This translation of the \lsti{env} field can be shared
with a derived family even if it extends \lsti{ty} (e.g., \lsti{STLCProd})---%
no recompilation is needed because \lsti{self.ty} is not tied to any
concrete definition of \lsti{ty}.

The type of \lsti{STLC\_env}'s \lsti{self} parameter is \lsti{STLC\_env\_Ctx},
a module type constructed from \lsti{STLC\_ty}, 
the translation of the field before \lsti{env},
and its context \lsti{STLC\_ty\_Ctx}.
In turn, \lsti{STLC\_ty\_Ctx} (not shown in \cref{fig:stlc-compiled}) is
constructed from \lsti{STLC\_subst},
the translation of the field before \lsti{ty},
and its context \lsti{STLC\_subst\_Ctx}.
Thus, the \lsti{self} parameter can be used to reference those and only
those fields in the current field's typing context, which echoes the
discussion in \cref{sec:sound}.

@@

An \lsti{FInductive} definition is translated to a parameterized module type.
Consider the inductive type \lsti{tm}.
In \lsti{STLC}, it is translated to a top-level module type
\lsti{STLC\_tm} that declares a \lsti{tm} type,
its constructors (e.g., \lsti{tm_abs}),
its partial recursor (\lsti{tm_prect_STLC}),
and the propositional equalities on the partial recursor
(e.g., \lsti{tm_abs_eq_STLC}; \cref{sec:prec}).

@@

The command \lsti{FInductive tm : Set += tm_fix : }\dadada~in family
\lsti{STLCFix} is again translated to a module type \lsti{STLCFix\_tm} (\cref{fig:stlcfix-compiled}).
It includes all the names declared by \lsti{STLC\_tm}
via command \lsti!Include STLC\_tm(self)!,
and additionally declares
the new constructor \lsti{tm_fix} and related equalities.

@@

Upon \lsti{End}~\lsti{subst}, a module type named \lsti{STLC\_subst} is generated.
%
As discussed in \cref{sec:latebind}, \lsti{subst} can be further bound,
so its definition is not exposed to the fields that come after it.
Accordingly, the translation \lsti{STLC\_subst} merely declares the type of \lsti{subst}
and the existence of its propositional equalities, leaving \lsti{subst} undefined
and the equalities unproven.
%
The propositional equalities are stated in terms of the case handlers,
whose definitions \emph{are} available through the \lsti{self} parameter.
So Coq can simplify, for example, \lsti{subst_tm_unit_eq}'s type to
\lsti!\forall x t, self.subst self.tm_unit x t = self.tm_unit!.
The equalities will be included and available for use in the
translations of the subsequent fields through their \lsti{self}
parameters.

Importantly, code generated for the case handlers is shared with derived families.
In \cref{fig:stlcfix-compiled}, module \lsti{STLCFix\_subst\_Cases}
reuses---without rechecking---all the case handlers in
\lsti{STLC\_subst\_Cases} via command \lsti{Include STLC\_subst\_Cases(self)}.

@@

In family \lsti{STLC},
field \lsti{env} and the case handlers for \lsti{subst}
are not further-bindable by derived families.
In contrast, \lsti{tm}, \lsti{subst}, and the related equalities
can be further bound.
The distinction is reflected in the translations.
The further-bindable fields are translated to module types that export only types of the fields.
The non-further-bindable fields are translated to modules that export definitional equalities
on the fields.
Opaque fields can be further bound, too (\cref{sec:override});
they are translated to modules that export opaque fields.

@@

Upon the conclusion of a family definition,
a representation of the family is created.
For example, module \lsti{STLC} in \cref{fig:stlc-compiled} is generated
upon \lsti{End}~\lsti{STLC}.
%
This module can be viewed as the ``fixed point'' of the
\lsti{self}-parameterized translations.
The ``fixed point'' is taken step by step, by adding the translation of
each field to this module in the same order as they appear in the family
definition.

For the non-further-bindables, the translated modules are directly included
(e.g., \lsti{Include STLC\_env} and \lsti{Include STLC\_subst\_Cases} in \cref{fig:stlc-compiled}).
%
The instantiation of \lsti{self}s for these modules is implicit, thanks to a Coq nicety:
when including a higher-order module, Coq automatically instantiates its
parameter with the current interactive module environment.
For instance, command \lsti{Include STLC\_subst\_Cases} is successfully executed,
because Coq automatically instantiates the \lsti{self} parameter using the current
module environment, which by construction contains all the fields required by \lsti{STLC\_subst\_Cases\_Ctx}.

For the further-bindables, \lsti{Axiom}s declared in the module types must be instantiated.
%
\begin{itemize}
  [align=left,itemsep=2pt,labelsep=*,leftmargin=*]

\item
In \cref{fig:stlc-compiled},
an inductive type \lsti{tm} is generated, instantiating the axiomatized
\lsti{tm} type and its constructors.
%
The partial recursor \lsti{tm_prect_STLC} is defined
%to \lsti{\lambda}\,\lsti{P, tm_rect (\lambda}\,\lsti{t, option (P t))},
with the help of \lsti{tm_rect}, the recursor Coq generates for \lsti{tm}.
Proving the equalities on \lsti{tm_prect_STLC} is then trivial, by \lsti{reflexivity}.

\item
Similarly, \lsti{subst} is instantiated by applying the recursor \lsti{tm_rect}
to the (already included) case handlers. %(\lsti{subst\_tm_unit}, etc.).
Propositional equalities on \lsti{subst} are then trivial, by \lsti{reflexivity}.

@@

Module \lsti{STLCFix} in \cref{fig:stlcfix-compiled} is emitted upon
\lsti{End}~\lsti{STLCFix}, in the same way as described above for \lsti{STLC}.
The translation makes sharing evident.
In particular, case handlers compiled for \lsti{STLC}
are reused to instantiate \lsti{subst}, \lsti{substlem}, and alike.
\lsti{STLC\_env} and \lsti{STLC\_typesafe}
are also reused in the construction of module \lsti{STLCFix}.
%
One may argue that since the first four constructors of \lsti{tm} are
repeated in \lsti{STLCFix}, the translation does not satisfy the
modular compilation requirement.
We could address this concern by using wrapper types,
but we consider restating constructors a small price to pay in return
for the clarity and concision of implementation.
We emphasize that compiled case handlers, such as \lsti!subst\_tm_abs!,
are entirely reusable without rechecking, even with restated constructors.
%
Finally, the reference \lsti{STLCFix.typesafe} (where \lsti{STLCFix} is a family)
can simply be translated to \lsti{STLCFix.typesafe} (where \lsti{STLCFix} is a Coq module),
as the last line of \cref{fig:stlcfix-compiled} shows.

### 2022-11-10 metatheory-condensed.tex
edits

W-types~\cite{martin1984intuitionistic} are a succinct way to model
inductive types in MLTT;
together with the identity type $\TyId{\cDot}{\cDot}$, they can express a
whole host of inductive types~\cite{hugunin2021not}.
%
A constructor is given by a pair of types
$\goodType{\Gm}{A}{}$ and $\goodType{\Gm,A}{B}{}$ (rule \ruleref{wsig/add}):
given two arguments $\goodTerm{\Gm}{t_1}{A}$ and
$\goodTerm{\Gm,\sub{B}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{T}$,
$\Wsup{\wsig}{t_1}{t_2}$ constructs a term of type~$T$ (rule \ruleref{tm/wsup}).
As the rules suggest, our formulation additionally supports
\emph{W-type signatures}.
A well formed W-type signature \goodWSig{\Gm}{\wsig}{n} is
composed of $n$ pairs of types, each of which represents a constructor
(rules \ruleref{wsig/empty} and \ruleref{wsig/add}),
with $\wsigproj{i}{1}{\wsig}$ and $\wsigproj{i}{2}{\wsig}$ giving the
$i$-th pair in~$\wsig$, and $\wcode{\wsig}$ giving the code of the W-type (\ruleref{tm/w}).

W-types are eliminated with the form $\Wrec{\wsig}{\lkg}{t}$, where $t$ is of a
W-type $\El{\wcode{\wsig}}$, and $\lkg$ is essentially an $n$-tuple of
case handlers for the $n$ constructors in~$\wsig$ (\ruleref{tm/wrec}).
Each case handler has a type of the form $\CaseSig{A}{B}{T}$, where
$T$ is the motive of the recursion (\ruleref{tyeq/casety}).
This collection of case handlers then encodes those defined or inherited
by an \lsti{FRecursion} command in our plugin.
We choose to type it with a linkage type $\TyLkg{\RecSig{\wsig}{T}}$
to avoid introducing $n$-tuples, which linkages generalize.

@@

In contrast, $\wsig$ should be hidden from the typing context of any
term that does not invoke $\Wsup{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$, so that
the term can be reused---without being rechecked---for a different
W-type signature~$\wsig'$ that extends~$\wsig$ with additional
constructors.
Moreover, the typing of the term should be made parametric to
the definitions of those fields that invoke $\texttt{Wsup}$ or $\texttt{Wrec}$,
so that the term can be reused---without being rechecked---when those fields
are overridden to support the extended signature~$\wsig'$.
%
Such abstraction required by family polymorphism is supported via linkages.

@@

In rule \ruleref{l/add}, the second premise $\goodTerm{\Gm, A}{t}{T}$ is
responsible for late binding.
Here, $A$ abstracts the fields preceding the current field~$t$
(i.e., those recorded by the first premise $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$).
Crucially, unlike the second field in a dependent pair (\ruleref{tm/pair}),
the typing of~$t$ in \ruleref{l/add} is polymorphic to how the fields in
$t$'s context are defined, so that a different linkage that overrides
fields in $t$'s context can reuse~$t$---without retyping it---by
projecting it (\ruleref{l/proj/2}) and then including it.

@@

\item 
The next four fields model the four constructors of~\lsti{tm}.
Constructor \lsti{tm_unit} is modeled as $\Wsup{\wsig_{\texttt{tm}}}{\top}{\bot}$
and typed with $\El{\mathit{self}.\texttt{tm}}$, where $\mathit{self}$ stands for
the typing context containing the first field~\lsti{tm}.
Because the W-type signature $\wsig_{\texttt{tm}}$ is exposed to this typing context,
$\El{\mathit{self}.\texttt{tm}}$ and $\El{\wcode{\wsig_{\texttt{tm}}}}$ can be equated,
as required by rule \ruleref{tm/wsup}.

\item
Similarly, $\wsig_{\texttt{tm}}$ is exposed to the typing context of
$t_{10}$, which models the recursive function \lsti{subst} by invoking
the recursor $\Wrec{\wsig_{\texttt{tm}}}{\cDot}{\cDot}$.
%
Like \lsti{subst}, partial recursors in \Lang are axiomatized by the plugin
(\cref{sec:latebind,sec:prec}), and they can similarly be defined in \TT
using $\texttt{Wrec}$.

\item
By contrast, $\wsig_{\texttt{tm}}$ is hidden from the typing context of all other fields.
Their typing depends on the knowledge that \lsti!tm! has type~$\cU$,
rather than \TyS{\wcode{\wsig_{\texttt{tm}}}},
so they can be reused in a context where \lsti!tm! is defined as $\wcode{\wsig'_{\texttt{tm}}}$,
where $\wsig'_{\texttt{tm}}$ extends~$\wsig_{\texttt{tm}}$ with additional constructors.
%
For example, in \cref{fig:stlc-linkage-typing}, the typing of the case handlers of
\lsti{subst} is agnostic of the definition of \lsti!tm!---it sees only $\text{\lsti!tm!}:\cU$---so
they can be reused by a linkage modeling \lsti{STLCFix}.

@@

Inheritance and code reuse can already be expressed through projection
of fields out of linkages and their inclusion into new linkages.
To make common patterns of linkage manipulations more convenient,
\TT provides a ``library'' of \emph{linkage transformers}, whose well-formedness judgments
have form $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$. The idea is that
applying~$h$ to a linkage of type~$\TyLkg{\lsig_1}$ yields a linkage of
type~$\TyLkg{\lsig_2}$ (\ruleref{tm/inh}).
%Linkage transformers are second-class; the only way to use them is to
%apply them to linkages to form new linkages.

Derived families can be modeled as linkage transformers inductively
constructed from the introduction forms
$\small\InhId$, $\small\InhExt{h}{t}$, $\small\InhOv{h}{t}$, $\small\InhInh{h}$, etc.
\cref{fig:fmltt-selected} shows the $\beta$-rule of an exemplary transformer,
\ruleref{tmeq/ov/beta}.
It states that applying the transformer $\small\InhOv{h}{t_2}$ to
a linkage of form $\LkgAdd{\lkg}{t_1}$
overrides the linkage's last field~$t_1$ with~$t_2$.
For instance, the construction below shows that 
$\small\InhOv{\InhId}{\wcode{\wsig'_{\texttt{tm}}}}$
is used as the first step in creating a linkage transformer modeling a derived family that
overrides $\wsig_{\texttt{tm}}$ with an extended
signature~$\wsig'_{\texttt{tm}}$.
\vspace{.9ex}

>>>>

W-types~\cite{martin1984intuitionistic} are a succinct way to model
inductive types in MLTT.
Together with the identity type $\TyId{\cDot}{\cDot}$, they can express a
whole host of inductive types~\cite{hugunin2021not}.
%
A constructor is given by a pair of types
$\goodType{\Gm}{A}{}$ and $\goodType{\Gm,A}{B}{}$ (rule \ruleref{wsig/add}):
given two arguments $\goodTerm{\Gm}{t_1}{A}$ and
$\goodTerm{\Gm,\sub{B}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{T}$,
$\Wsup{\wsig}{t_1}{t_2}$ constructs a term of type~$T$ (rule \ruleref{tm/wsup}).
As the rules suggest, our formulation additionally supports
\emph{W-type signatures}.
A well formed W-type signature \goodWSig{\Gm}{\wsig}{n} is
composed of $n$ pairs of types, each of which represents a constructor
(rules \ruleref{wsig/empty} and \ruleref{wsig/add}),
with $\wsigproj{i}{1}{\wsig}$ and $\wsigproj{i}{2}{\wsig}$ giving the
$i$-th pair in~$\wsig$, and $\wcode{\wsig}$ giving the code of the W-type (\ruleref{tm/w}).

W-types are eliminated with the form $\Wrec{\wsig}{\lkg}{t}$, where $t$ is of a
W-type $\El{\wcode{\wsig}}$, and $\lkg$ is essentially an $n$-tuple of
case handlers for the $n$ constructors in~$\wsig$ (\ruleref{tm/wrec}).
Each case handler has a type of the form $\CaseSig{A}{B}{T}$, where
$T$ is the motive of the recursion (\ruleref{tyeq/casety}).
For simplicity, we model non-dependent motives.
This collection of case handlers~$\lkg$ then encodes those defined and inherited
by an \lsti{FRecursion} command in our plugin.
We choose to type it with a linkage type $\TyLkg{\RecSig{\wsig}{T}}$
to avoid introducing non-dependent $n$-tuples, which linkages generalize.

@@

In contrast, $\wsig$ should be hidden from the typing context of any
term that does not invoke $\Wsup{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$, so that
the term can be reused---without being rechecked---for a different
W-type signature~$\wsig'$ that extends~$\wsig$ with additional
constructors.
Moreover, the typing of the term should be made parametric to
the definitions of those fields that invoke $\texttt{Wsup}$ or $\texttt{Wrec}$,
so that the term can be reused---without being rechecked---when those fields
are overridden to support the extended signature~$\wsig'$.
%
Such abstraction required by family polymorphism is supported via linkages,
which we discuss next.

@@

In rule \ruleref{l/add}, the second premise $\goodTerm{\Gm, A}{t}{T}$ is
responsible for late binding.
Here, $A$ abstracts the fields preceding the current field~$t$
(i.e., those recorded by the first premise $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$).
Crucially, unlike the second field in a dependent pair (\ruleref{tm/pair}),
the typing of~$t$ in \ruleref{l/add} is polymorphic to how the fields in
$t$'s context are defined, so that a different linkage that overrides
fields in $t$'s context can reuse~$t$---without retyping it---by
projecting it and then including it.

@@

\item 
The next four fields model the four constructors of~\lsti{tm}.
Constructor \lsti{tm_unit} is modeled as $\Wsup{\wsig_{\texttt{tm}}}{\top}{\bot}$
and typed with $\El{\mathit{self}.\texttt{tm}}$, where $\mathit{self}$ stands for
the typing context containing the first field~\lsti{tm}.
The W-type signature $\wsig_{\texttt{tm}}$ is exposed in
this typing context; that is, \lsti{tm} has type $\TyS{\wcode{\wsig_{\texttt{tm}}}}$.
So $\El{\mathit{self}.\texttt{tm}}$ and $\El{\wcode{\wsig_{\texttt{tm}}}}$ can be equated,
as required by rule \ruleref{tm/wsup}.

\item
Similarly, $\wsig_{\texttt{tm}}$ is exposed in the typing context of
$t_{10}$, which models the recursive function \lsti{subst} by invoking
the recursor $\Wrec{\wsig_{\texttt{tm}}}{\cDot}{\cDot}$.
%
Like \lsti{subst}, partial recursors in \Lang are axiomatized by the plugin
(\cref{sec:latebind,sec:prec}), and they can similarly be defined in \TT
using $\texttt{Wrec}$.

\item
By contrast, $\wsig_{\texttt{tm}}$ is hidden from the typing of all other fields.
Their typing depends on the knowledge that \lsti!tm! has type~$\cU$,
rather than \TyS{\wcode{\wsig_{\texttt{tm}}}},
so they can be reused in a context where \lsti!tm! is defined as $\wcode{\wsig'_{\texttt{tm}}}$,
where $\wsig'_{\texttt{tm}}$ extends~$\wsig_{\texttt{tm}}$ with additional constructors.
%
For example, in \cref{fig:stlc-linkage-typing}, the typing of the case handlers of
\lsti{subst} (e.g., $t_6$) is oblivious to the definition of \lsti!tm!---it sees only $\text{\lsti!tm!}:\cU$---so
the case handlers can be reused by a linkage modeling \lsti{STLCFix}.

@@

Inheritance and code reuse can already be expressed through the projection
of fields out of linkages and their inclusion into new linkages.
To make common patterns of linkage manipulations more convenient,
\TT provides a ``library'' of \emph{linkage transformers}, whose well-formedness judgments
have form $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$. The idea is that
applying~$h$ to a linkage of type~$\TyLkg{\lsig_1}$ yields a linkage of
type~$\TyLkg{\lsig_2}$ (\ruleref{tm/inh}).
%Linkage transformers are second-class; the only way to use them is to
%apply them to linkages to form new linkages.

Derived families can be modeled as linkage transformers inductively
constructed from the introduction forms
$\small\InhId$, $\small\InhExt{h}{t}$, $\small\InhOv{h}{t}$, $\small\InhInh{h}$, etc.
\cref{fig:fmltt-selected} shows the $\beta$-rule of an exemplary transformer,
\ruleref{tmeq/ov/beta}.
It states that applying the transformer $\small\InhOv{h}{t_2}$ to
a linkage of form $\LkgAdd{\lkg}{t_1}$
overrides the linkage's last field~$t_1$ with~$t_2$.
For instance, the construction below shows that 
$\small\InhOv{\InhId}{\wcode{\wsig'_{\texttt{tm}}}}$
is used as the first step in creating a linkage transformer modeling a derived family that
overrides $\wsig_{\texttt{tm}}$ with an extended
signature~$\wsig'_{\texttt{tm}}$:
\vspace{.9ex}
